### 2022-05-03 response_final.tex

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}

\usepackage{color}

\usepackage{xspace}

\usepackage{lmodern}

\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}

\usepackage[pdfencoding=auto, psdextra]{hyperref}

\usepackage{natbib}
\bibliographystyle{chicago}

\newcommand{\eref}[1]{Eq.~(\ref{eq:#1})}
\newcommand{\fref}[1]{Fig.~\ref{fig:#1}}

%% Consistent, changeable style for subscripts
\newcommand{\vvvar}{\mathrm{var}}
\newcommand{\wwwt}{\mathrm{wt}}

\newcommand{\rx}[1]{\ensuremath{{r}_{#1}}\xspace} 
\newcommand{\ry}[1]{\rx{\mathrm{#1}}} 
\newcommand{\rw}{\rx{\wwwt}}
\newcommand{\rv}{\rx{\vvvar}}

\newcommand{\Rx}[1]{\ensuremath{{\mathcal R}_{#1}}\xspace} 
\newcommand{\Ry}[1]{\Rx{\mathrm{#1}}}
\newcommand{\Ro}{\Rx{0}}
\newcommand{\RR}{\ensuremath{{\mathcal R}}\xspace}
\newcommand{\Rw}{\Rx{\wwwt}}
\newcommand{\Rv}{\Rx{\vvvar}}

\newcommand{\dd}[1]{\ensuremath{\, \mathrm{d}#1}}
\newcommand{\dtau}{\dd{\tau}}
\newcommand{\dx}{\dd{x}}
\newcommand{\dsigma}{\dd{\sigma}}

\newcommand{\rev}{\subsection*}
\newcommand{\revtext}{\textsf}
\setlength{\parskip}{\baselineskip}
\setlength{\parindent}{0em}

\newcommand{\comment}[3]{\textcolor{#1}{\textbf{[#2: }\textsl{#3}\textbf{]}}}
\newcommand{\jd}[1]{\comment{cyan}{JD}{#1}}
\newcommand{\swp}[1]{\comment{magenta}{SWP}{#1}}
\newcommand{\dc}[1]{\comment{blue}{DC}{#1}}
\newcommand{\jsw}[1]{\comment{green}{JSW}{#1}}
\newcommand{\hotcomment}[1]{\comment{red}{HOT}{#1}}

\newcommand{\psymp}{\ensuremath{p}} %% primary symptom time
\newcommand{\ssymp}{\ensuremath{s}} %% secondary symptom time
\newcommand{\pinf}{\ensuremath{\alpha_1}} %% primary infection time
\newcommand{\sinf}{\ensuremath{\alpha_2}} %% secondary infection time

\newcommand{\psize}{{\mathcal P}} %% primary cohort size
\newcommand{\ssize}{{\mathcal S}} %% secondary cohort size

\newcommand{\gtime}{\tau_{\rm g}} %% generation interval
\newcommand{\gdist}{g} %% generation-interval distribution
\newcommand{\idist}{\ell} %% incubation period distribution

\newcommand{\total}{{\mathcal T}} %% total number of serial intervals

\begin{document}

\noindent Dear Editors:

We are submitting the final version our manuscript, “The importance of the generation interval in investigating dynamics and control of new SARS-CoV-2 variants”.

Thank you for your consideration of our submission.

Sincerely,

Sang Woo Park

\rev{Reviewer \#1}

\revtext{When I started reviewing this paper, I was highly skeptical of the notions of strength and speed for epidemics and interventions. This paper has convinced me that these are actually very useful concepts, certainly in theory if not necessarily in practice. The assumptions are mostly clearly explained, but the assumption of homogeous mass-action should be made explicit early in the paper (though many of the qualitative insights probably transfer beyond this class of models). While the results are shown in only a special case (homogeneous massaction with a transmission kernal that is constant in t and a gamma-distributed generation interval), it is an important special case because it is similar to many models used in practice. The discussion clearly addresses situations where the strength/speed dichotomy and several assumptions might break down, but it makes a good case that the qualitative insights are still useful. I have reviewed very few papers that changed my mind about something so important, and I expect this will be equally useful to many other readers. My only general comment is that the presentation of the basic insights and the figures sometimes seem scattered; I found myself jumping around quite a bit while reviewing the paper. I have a few more specific comments below.}

Thank you for your review. We believe that a part of the scattered-ness is caused by pdf conversion process from a latex document---for example, figure 3 is described in section 5 (pages 10--11) but appears in page 12 due to its size. Likewise, figures 4 and 5 have been pushed back to the end of section 6. We think that these issues will be less problematic when the paper is properly formatted according to the journal guideline. Otherwise, we have made minor changes throughout the manuscript to try to tie the results closer to the figures whenever appropriate and re-ordered some text for clearer presentation.

\revtext{(page 3, lines 43–44) It is not true that R determines the final size of an epidemic in a homogeneous population. For example, the final size of an epidemic in a homogeneous population on a configuration model network depends on the degree distribution as well as $\RR$, and this final size corresponds to the mass-action final size with the same R only when the degree distribution is Poisson. It would be accurate to say that $\RR$ determines the final size of an epidemic in a homogeneous mass-action model. The entire paper is based on homogeneous mass-action models, so it help to make this explicit.}

We have reworded this sentence as follows:

``The epidemic strength also determines the final size of an epidemic in a homogeneously mixing population under the mass-action assumption''

\revtext{(page 4, lines 90 and 92) ``$K_{\mathrm{post}}$'' should be ``$K_{\mathrm{post}}(\tau)$''}

Done.

\revtext{(page 4, lines 93–94) The ``kernel'' needs to be explained more clearly before the statement that intervention speed ($\phi$) must be greater than epidemic speed ($r$) in order to control the epidemic. The explanation given could be interpreted to mean that the pre-intervention generation interval distribution $g_x(\tau)$ should be replaced by something proportional
to $g_x(\tau) \exp(-\phi \tau)$. If you then imitate the derivation from Wallinga and Lipsitch (American Journal of Epidemiology, 2007) using the pre- and post-intervention kernels, you get $\RR_{\textrm{pre}} = M_g(-r)^{-1}$ and $$\RR_{\textrm{post}} = \frac{M_g(-\phi)}{M_g(-r-\phi)}$$
where $M_g()$ is the moment generating function for the PDF $g(\tau)$. This
does not lead to $\RR_{\textrm{post}} < 1$ if $\phi > r$. However, we do get $\RR_{\textrm{post}} < 1$ if $\phi < -r$, which would seem to correspond to an intervention that makes people more infectious over time. 
This counterintuitive result comes from the need to re-normalize $g(\tau) e^{-\phi \tau}$.
In order to get a PDF that integrates to one, and it shows the importance of clearly maintaining the distinction between the generation interval distribution and the transmission kernel.}

Thank you for pointing this out. We agree that making the distinction between the generation-interval distribution and the transmission kernel is important. We have rephrased this paragraph to try to make the distinction clearer.

``While both approaches are reasonable, holding different quantities constant (i.e., strength or speed) can lead to different conclusions about the spread of the pathogen and its control \citep{doi:10.1098/rspb.2020.1556}.
To illustrate the differences in conclusions when holding $\RR$ or $r$ fixed, we consider two idealized interventions of constant strength and constant speed.
Before these interventions are introduced, the dynamics of pathogen spread can be characterized in terms of the pre-intervention kernel $K_{\mathrm{pre}}(\tau)$, which represents the rate at which an infected individual generates secondary infections $\tau$ time units after infection.
The pre-intervention kernel can be further decomposed in terms of the epidemic strength $\RR$ and the generation-interval distribution $g(\tau)$: $K_{\mathrm{pre}}(\tau) = \RR g(\tau)$;
therefore, the infection knerel integrates to $\RR$ while the generation-interval distribution integrates to 1.
Then, a constant-strength intervention reduces transmission by a constant factor $\theta$ throughout infection such that the post-intervention kernel is $K_{\mathrm{post}}(\tau) = K_{\mathrm{pre}}(\tau)/\theta$.
In this case, the intervention strength $\theta$ must be greater than the epidemic strength $\RR$ to control the epidemic.
In contrast, a constant-speed intervention reduces transmission after infection by a constant rate $\phi$ throughout infection: $K_{\mathrm{post}}(\tau) = K_{\mathrm{pre}}(\tau) \exp(-\phi \tau)$.
In this case, the intervention speed $\phi$ must be greater than the pre-intervention epidemic speed $r$ to control the epidemic.
We note that the resulting post-intervention generation-interval distribution under a constant-speed intervention is \emph{not} equal to $g(\tau) \exp(-\phi \tau)$; instead $g(\tau) \exp(-\phi \tau)$ needs to be renormalized to integrate to 1 because the generation-interval distribution is a probability distribution.''

\revtext{(page 6, line 192) It would be helpful to point out that $\kappa^{-1}$ is the gamma shape parameter.}

Done.

\revtext{(page 9, lines 270–277) It would help to explain that Rwt = 0.49 corresponds to the top left of Figure 2 and Rwt = 1.27 corresponds to the bottom right.}

Done.

\revtext{(page 10, lines 282–290) I appreciate the careful distinction made between
the case reproduction number and the instantaneous reproduction number
as well as the explanation of why the instantaneous reproduction is the
more relevant here}

Thank you.

\revtext{(page 14, Figure 4) Subfigures A, C, E and B, D, F should have a label
for the x-axis. To make room, the y-axis labels in G and H could be just
Rvar and rvar, respectively.}

We have added axis titles for the subfigures.

\revtext{(page 19, lines 513–529) In practice, generation intervals can be observed
``forwards'' and ``backwards'' but both are subject to bias (as explained in
Ref. 19). Do the authors have an opinion about which of these would allow
a more reliable estimate of the relative length of the generation intervals
of a variant compared to wild type?}

We have added a new paragraph:

``While detailed contact tracing data can help narrow down uncertainties in the generation-interval distributions, there are additional complexities to comparing generation-interval estimates of different variants.
The generation intervals can be measured in two different ways: backward and forward \citep{champredon2015intrinsic}.
The backward distribution starts from the cohort of infectees who were infected at the same time and looks at when their infectors were infected. 
The backward distribution is subject to dynamical biases because we are more likely to observe recent infections (and therefore shorter generation intervals) when the epidemic is growing.
On the other hand, the forward distribution starts from the cohort of infectors who were infected at the same time and looks at when they transmitted infections.
While the forward distribution is more reliable, changes in epidemic dynamics due to intervention measures and susceptible depletions can affect the shape of the distribution \citep{ali2020serial, nishiura2010time}.
For example, comparing generation-interval estimates from different intervention periods will necessarily bias the true differences in the generation-interval distributions of different variants.''

\bibliography{newvariant_abbv.bib}

\end{document}

### 2022-05-03 newvariant.tex
really done

Estimating variant epidemic strength and speed remains a key question in understanding the threat of SARS-CoV-2 variants of concern (VoCs) \cite{switzerland2021variant,davies2021estimated,di2021impact,graham2021changes,leung2021early,volz2021transmission,zhao2021,campbell2021increased}.
Epidemic ``strength'' is measured by the reproduction number $\RR$---a unitless quantity representing the average number of new infections caused by a typical infection. 
A pathogen can spread in a population if $\RR>1$ \citep{diekmann1990definition}.
The epidemic strength also determines the final size of an epidemic in a homogeneous population \citep{anderson1991infectious}.
Epidemic ``speed'' is characterized by the growth rate $r$, which has units of $1/\mathrm{time}$ and describes the exponential rate of pathogen spread at the population level.
Like epidemic strength, epidemic speed also determines conditions for pathogen elimination: $r=0$ is a threshold equivalent to $\RR=1$ under constant conditions.
However, epidemiological modelers have often over-emphasized $\RR$ at the expense of $r$ \citep{doi:10.1098/rspb.2020.1556};
we thus use the terms ``strength'' and ``speed''  here to underline our contention that these metrics are better seen as complementary perspectives (and to link them to complementary perspectives on measuring the transmission advantage of new variants).

@@

While both approaches are reasonable, holding different quantities constant (i.e., strength or speed) can lead to different conclusions about the spread of the pathogen and its control \citep{doi:10.1098/rspb.2020.1556}.
To illustrate the differences in conclusions when holding $\RR$ or $r$ fixed, we consider two idealized interventions of constant strength and constant speed.
Given a pre-intervention kernel $K_{\mathrm{pre}}(\tau) = \RR g(\tau)$, which represents the rate at which an infected individual generates secondary infections $\tau$ time units after infection, a constant-strength intervention reduces transmission by a constant factor $\theta$ across generation intervals such that the post-intervention kernel is $K_{\mathrm{post}} = K_{\mathrm{pre}}(\tau)/\theta$.
In this case, the intervention strength $\theta$ must be greater than the epidemic strength $\RR$ to control the epidemic.
In contrast, a constant-speed intervention reduces transmission after infection by a constant rate $\phi$ across generation intervals: $K_{\mathrm{post}} = K_{\mathrm{pre}}(\tau) \exp(-\phi \tau)$.
In this case, the intervention speed $\phi$ must be greater than the epidemic speed $r$ to control the epidemic.

@@

Over a short period of time, we can assume that epidemiological conditions remain roughly constant: $\RR_x(t) \approx \RR_x$ and $g_x(t, \tau) \approx g_x(\tau)$, in which case the incidence of infections changes exponentially.
Here, we use the term ``epidemiological conditions'' to broadly refer to all factors that affect transmission at the population-level---mathematically, they are captured by the kernel $K(t, \tau)$.
In the context of SARS-CoV-2 infections, we are essentially assuming that the changes in susceptible pool, behavior, and contact rates are usually small over a short period of time---this assumption would not apply at the moment of a drastic policy change, but could be applied to the periods immediately before and after.
Then, the incidence of each strain grows (or decays) exponentially at rate $r_x$, satisfying the Euler-Lotka equation \citep{wallinga2007generation}:
\begin{equation}
\frac{1}{\RR_x} = \int_0^\infty \exp(- r_x \tau) g_x(\tau) \dtau.
\end{equation}
We can approximate this $r$--$\RR$ relationship by assuming that the generation-interval distribution is Gamma-distributed, and summarizing it using the mean generation interval $\bar{G}_x$ and the squared coefficient of variation $\kappa_x$ \citep{park2019practical}:
\begin{equation}
\RR_x \approx (1 + \kappa_x r_x \bar{G}_x)^{1/\kappa_x}.
\end{equation}
Various Gamma-generation-interval assumptions have been widely used in epidemic modeling, including for models of SARS-CoV-2 \citep{doi:10.1098/rsif.2020.0144}.
The Gamma-generation-interval assumption includes as a special case models that assume exponentially distributed generation intervals (when $\kappa=1$), corresponding to the SIR model \citep{anderson1991infectious}.
We note that when the infectious periods are Gamma distributed---another standard assumption in epidemic modeling---the resulting generation interval does not necessarily follow the Gamma distribution (see \cite{roberts2007model} for detailed discussion).

@@

% We thus ask: what factors affect the relative strength $\rho = \Rv/\Rw$ of a new variant, conditional on an observed relative speed $\delta$?
Inference of the relative strength $\rho = \Rv/\Rw$ from the relative speed $\delta$ depends on assumptions about the generation-interval distributions of both strains.
Given the mean generation interval of the variant $\Gv$ and the wild type $\Gw$, the relative strength $\rho = \Rv/\Rw$ under the Gamma assumption \citep{park2019practical} is given by:
\begin{equation}
\rho = \left(\frac{1 + \kappa (\rw + \delta) \Gv}{1 + \kappa \rw \Gw}\right)^{1/\kappa}.
\end{equation}
Therefore, the relative strength $\rho$ depends not only on the relative speed $\delta$ and the generation-interval distributions but also on how fast the wild type is spreading in the population (\rw)---
some analyses have implicitly or explicitly neglected this factor by either assuming $\rw = 0$ \citep{switzerland2021variant} or $\kappa = 0$ \citep{davies2021estimated} (in the latter case, $\rho = \exp(\delta \Gw)$ when $\Gv=\Gw$).

@@

Unsurprisingly, we find that an increased speed of $\delta=0.1\pday$ for the variant is consistent with higher strength than the wild type across the range of epidemiological conditions considered (\fref{relstrength}).
However, the magnitude of relative strength $\rho$ is sensitive to assumptions about generation intervals:
For realistic values of $\kappa$ (excluding 0 and 1), the inferred relative strength $\rho$ ranges between 1.1 and 2.3 when $\Gv$ is allowed to vary between $2/3$ and $3/2$ of \Gw.

In general, longer mean generation intervals of the new variant translate to higher values of $\rho$ (and vice versa), except when $\rv \leq 0$ (recall, we always assume $\rw<\rv$).
When $\rv = 0$, we always have $\Rv = 1$ and so $\rho$ is independent of the generation-interval distribution of the new variant.
When $\rv < 0$, we see that longer generation intervals decrease $\rho$ because longer generation intervals actually lead to slower decay (higher $r$).
Assuming a narrower distribution for both the variant and the wild type strain (lower $\kappa$) has qualitatively similar effects as assuming longer generation intervals (leading to higher values of $\rho$ when $\rv > 0$ and lower values of $\rho$ when $\rv < 0$) because both reduce the amount of early transmission.
When $\rw < 0 < \rv$, inference of $\rho$ is relatively insensitive to values of $\kappa$.

@@

In general, longer generation intervals lead to slower relative speed of the variant when the incidence of both strains is increasing (\fref{relspeed}, bottom panels) because slower growth of the variant reduces the differences in absolute speed.
When $\Rv=1$, the relative speed is insensitive to the generation-interval distribution of the variant because we always have $\rv=0$.
When $\Rv<1$, longer generation intervals of the variant lead to slower decay ($\rv$ closer to 0), and therefore, greater relative speed.
We see that assuming a narrower distribution for both the variant and the wild type strain (lower $\kappa$) has qualitatively similar effects as assuming a longer mean (leading to lower values of $\delta$ when $\Rv > 0$ and higher values of $\delta$ when $\Rv < 0$).

\fref{relspeed} also shows that when $\rho$ is fixed, relative speed depends on underlying epidemiological conditions---specifically, the absolute strength of the two strains. 
For example, even when the generation-interval distributions are identical ($\Gv=\Gw$, in this case), changing $\Rw$ from 0.49 to 1.27 (and thus $\Rv$ from 0.79 to 2.04) changes the relative speed $\delta$ from 0.08 to 0.11 when $\kappa=0.2$---and from 0.06 to 0.14 when $\kappa=0.5$.
Differences in the generation-interval distributions exaggerate these changes.
Therefore, characterizing changes in the proportion of variants by assuming a fixed relative speed (e.g., by fitting a standard logistic growth curve \citep{switzerland2021variant} and using the resulting value of $\delta$) should be done with care.

@@

Instead of inferring relative strength from relative speed, one can separately estimate time-varying reproduction numbers $\RR(t)$ of the variant and the wild type from observed incidence and assumptions about the generation intervals, and then calculate the ratio---such methods have been used in previous analyses of the Alpha \citep{volz2021transmission} and Delta variants \citep{abbott2021estimating}.
Broadly, there are two types of time-varying reproduction numbers: case reproduction number and instantaneous reproduction number.
The case reproduction number is defined as the average number of secondary infections caused by an individual infected at time $t$ and therefore depends on transmission after time $t$ \citep{wallinga2004different}.
The instantaneous reproduction number is defined as the average number of secondary infections that would be caused by an individual infected at time $t$ if conditions were to remain the same \citep{fraser2007estimating}; 
therefore, the instantaneous reproduction number only depends on transmission at time $t$ and is most appropriate for real-time evaluation of changes in transmission \citep{gostic2020practical}.
The estimation of the instantaneous reproduction number was popularized by \cite{cori2013new} and has been widely adopted in epidemiological analyses of SARS-CoV-2 \citep{abbott2020estimating,knight2020estimating,flaxman2020Rt,brauner2021inferring,li2021temporal}.

@@

When the assumed distribution matches the true distribution, the estimated reproduction numbers match the true values (\fref{Rtbias}D); thus, their ratio remains constant and $\hat{\RR}_{\textrm{var}}(t)=\Rv(t)$ (\fref{Rtbias}E).
However, when the generation-interval distribution of the variant differs from the assumed distribution, the ratio changes over time (\fref{Rtbias}B,H).
If the true generation intervals of the variant have a shorter mean than the assumed distribution, we over-estimate $\Rv(t)$ during the growth phase and under-estimate $\Rv(t)$ during the decay phase (and conversely, \fref{Rtbias}A,G), which further translates to biases in the estimated relative strength (\fref{Rtbias}B,H).

In practice, estimates of instantaneous reproduction numbers $\RR(t)$ (and therefore, their ratios) can be noisy due to limited data availability or model assumptions;
instead, we might want to estimate a single value of relative strength $\rho$.
For example, we can estimate $\rho$ by plotting the estimated strength of the variant $\hat{\RR}_{\textrm{var}}(t)$ against the estimated strength of the wild type $\hat{\RR}_{\textrm{wt}}(t)$---as presented in Figure 2 of \cite{volz2021transmission}---and performing a linear regression (\fref{Rtbias}C,F,I).
Here, we estimate the slope while fixing the intercept to zero: the regression line should go through the origin in theory because $\Rv = 0$ when $\Rw = 0$.
If $\rho$ is constant, and generation-interval distributions are correctly specified, we obtain a straight line with a slope of $\rho$ and intercept at zero (\fref{Rtbias}F).
However, when the assumed mean generation interval is longer than the that of the variant, we over-estimate the slope (and conversely, \fref{Rtbias}C,I).

@@

>>>>

Estimating variant epidemic strength and speed remains a key question in understanding the threat of SARS-CoV-2 variants of concern (VoCs) \cite{switzerland2021variant,davies2021estimated,di2021impact,graham2021changes,leung2021early,volz2021transmission,zhao2021,campbell2021increased}.
Epidemic ``strength'' is measured by the reproduction number $\RR$---a unitless quantity representing the average number of new infections caused by a typical infection. 
A pathogen can spread in a population if $\RR>1$ \citep{diekmann1990definition}.
The epidemic strength also determines the final size of an epidemic in a homogeneously mixing population under the mass-action assumption \citep{anderson1991infectious}.
Epidemic ``speed'' is characterized by the growth rate $r$, which has units of $1/\mathrm{time}$ and describes the exponential rate of pathogen spread at the population level.
Like epidemic strength, epidemic speed also determines conditions for pathogen elimination: $r=0$ is a threshold equivalent to $\RR=1$ under constant conditions.
However, epidemiological modelers have often over-emphasized $\RR$ at the expense of $r$ \citep{doi:10.1098/rspb.2020.1556};
we thus use the terms ``strength'' and ``speed''  here to underline our contention that these metrics are better seen as complementary perspectives (and to link them to complementary perspectives on measuring the transmission advantage of new variants).

@@

While both approaches are reasonable, holding different quantities constant (i.e., strength or speed) can lead to different conclusions about the spread of the pathogen and its control \citep{doi:10.1098/rspb.2020.1556}.
To illustrate the differences in conclusions when holding $\RR$ or $r$ fixed, we consider two idealized interventions of constant strength and constant speed.
Before these interventions are introduced, the dynamics of pathogen spread can be characterized in terms of the pre-intervention kernel $K_{\mathrm{pre}}(\tau)$, which represents the rate at which an infected individual generates secondary infections $\tau$ time units after infection.
The pre-intervention kernel can be further decomposed in terms of the epidemic strength $\RR$ and the generation-interval distribution $g(\tau)$: $K_{\mathrm{pre}}(\tau) = \RR g(\tau)$;
therefore, the infection kernel integrates to $\RR$ while the generation-interval distribution integrates to 1.
Then, a constant-strength intervention reduces transmission by a constant factor $\theta$ throughout infection such that the post-intervention kernel is $K_{\mathrm{post}}(\tau) = K_{\mathrm{pre}}(\tau)/\theta$.
In this case, the intervention strength $\theta$ must be greater than the epidemic strength $\RR$ to control the epidemic.
In contrast, a constant-speed intervention reduces transmission after infection by a constant rate $\phi$ throughout infection: $K_{\mathrm{post}}(\tau) = K_{\mathrm{pre}}(\tau) \exp(-\phi \tau)$.
In this case, the intervention speed $\phi$ must be greater than the pre-intervention epidemic speed $r$ to control the epidemic.
We note that the resulting post-intervention generation-interval distribution under a constant-speed intervention is \emph{not} equal to $g(\tau) \exp(-\phi \tau)$; instead $g(\tau) \exp(-\phi \tau)$ needs to be renormalized to integrate to 1 because the generation-interval distribution is a probability distribution.

Given a pre-intervention kernel $K_{\mathrm{pre}}(\tau) = \RR g(\tau)$, which represents the rate at which an infected individual generates secondary infections $\tau$ time units after infection, a constant-strength intervention reduces transmission by a constant factor $\theta$ across generation intervals such that the post-intervention kernel is $K_{\mathrm{post}}(\tau) = K_{\mathrm{pre}}(\tau)/\theta$.
In this case, the intervention strength $\theta$ must be greater than the epidemic strength $\RR$ to control the epidemic.
In contrast, a constant-speed intervention reduces transmission after infection by a constant rate $\phi$ across generation intervals: $K_{\mathrm{post}}(\tau) = K_{\mathrm{pre}}(\tau) \exp(-\phi \tau)$.
In this case, the intervention speed $\phi$ must be greater than the epidemic speed $r$ to control the epidemic.

@@

Over a short period of time, we can assume that epidemiological conditions remain roughly constant: $\RR_x(t) \approx \RR_x$ and $g_x(t, \tau) \approx g_x(\tau)$, in which case the incidence of infections changes exponentially.
Here, we use the term ``epidemiological conditions'' to broadly refer to all factors that affect transmission at the population-level---mathematically, they are captured by the kernel $K(t, \tau)$.
In the context of SARS-CoV-2 infections, we are essentially assuming that the changes in susceptible pool, behavior, and contact rates are usually small over a short period of time---this assumption would not apply at the moment of a drastic policy change, but could be applied to the periods immediately before and after.
Then, the incidence of each strain grows (or decays) exponentially at rate $r_x$, satisfying the Euler-Lotka equation \citep{wallinga2007generation}:
\begin{equation}
\frac{1}{\RR_x} = \int_0^\infty \exp(- r_x \tau) g_x(\tau) \dtau.
\end{equation}
We can approximate this $r$--$\RR$ relationship by assuming that the generation-interval distribution is Gamma-distributed, and summarizing it using the mean generation interval $\bar{G}_x$ and the squared coefficient of variation $\kappa_x$ \citep{park2019practical}:
\begin{equation}
\RR_x \approx (1 + \kappa_x r_x \bar{G}_x)^{1/\kappa_x}.
\end{equation}
We note that $\kappa_x^{-1}$ also corresponds to the shape parameter of a Gamma distribution.
Various Gamma-generation-interval assumptions have been widely used in epidemic modeling, including for models of SARS-CoV-2 \citep{doi:10.1098/rsif.2020.0144}.
The Gamma-generation-interval assumption includes as a special case models that assume exponentially distributed generation intervals (when $\kappa=1$), corresponding to the SIR model \citep{anderson1991infectious}.
We note that when the infectious periods are Gamma distributed---another standard assumption in epidemic modeling---the resulting generation interval does not necessarily follow the Gamma distribution (see \cite{roberts2007model} for detailed discussion).

@@

We thus ask: what factors affect the relative strength $\rho = \Rv/\Rw$ of a new variant, conditional on an observed relative speed $\delta$?
Inference of the relative strength $\rho = \Rv/\Rw$ from the relative speed $\delta$ depends on assumptions about the generation-interval distributions of both strains.
Given the mean generation interval of the variant $\Gv$ and the wild type $\Gw$, the relative strength $\rho = \Rv/\Rw$ under the Gamma assumption \citep{park2019practical} is given by:
\begin{equation}
\rho = \left(\frac{1 + \kappa (\rw + \delta) \Gv}{1 + \kappa \rw \Gw}\right)^{1/\kappa}.
\end{equation}
Therefore, the relative strength $\rho$ depends not only on the relative speed $\delta$ and the generation-interval distributions but also on how fast the wild type is spreading in the population (\rw)---
some analyses have implicitly or explicitly neglected this factor by either assuming $\rw = 0$ \citep{switzerland2021variant} or $\kappa = 0$ \citep{davies2021estimated} (in the latter case, $\rho = \exp(\delta \Gw)$ when $\Gv=\Gw$).

@@

Unsurprisingly, we find that an increased speed of $\delta=0.1\pday$ for the variant is consistent with higher strength ($\rho > 1$) than the wild type across all five epidemiological scenarios considered (\fref{relstrength}).
However, the magnitude of relative strength $\rho$ is sensitive to assumptions about generation intervals:
For realistic values of $\kappa$ (excluding 0 and 1), the inferred relative strength $\rho$ ranges between 1.1 and 2.3 (across all five scenarios) when $\Gv$ is allowed to vary between $2/3$ and $3/2$ of \Gw.

In general, longer mean generation intervals of the new variant translate to higher values of $\rho$ (and vice versa;  \fref{relstrength}, bottom panels), except when $\rv \leq 0$ (recall, we always assume $\rw<\rv$; \fref{relstrength}, top panels).
When $\rv = 0$, we always have $\Rv = 1$ and so $\rho$ is independent of the generation-interval distribution of the new variant.
When $\rv < 0$, we see that longer generation intervals decrease $\rho$ because longer generation intervals actually lead to slower decay (higher $r$).
Assuming a narrower distribution for both the variant and the wild type strain (lower $\kappa$) has qualitatively similar effects as assuming longer generation intervals (leading to higher values of $\rho$ when $\rv > 0$ and lower values of $\rho$ when $\rv < 0$) because both reduce the amount of early transmission.
When $\rw < 0 < \rv$, inference of $\rho$ is relatively insensitive to values of $\kappa$.

@@

In general, longer generation intervals lead to slower relative speed of the variant when the incidence of both strains is increasing (\fref{relspeed}, bottom panels) because slower growth of the variant reduces the differences in absolute speed.
When $\Rv=1$, the relative speed is insensitive to the generation-interval distribution of the variant because we always have $\rv=0$ (\fref{relspeed}, top-right panel).
When $\Rv<1$, longer generation intervals of the variant lead to slower decay ($\rv$ closer to 0), and therefore, greater relative speed (\fref{relspeed}, top-left panel).
We see that assuming a narrower distribution for both the variant and the wild type strain (lower $\kappa$) has qualitatively similar effects as assuming a longer mean (leading to lower values of $\delta$ when $\Rv > 0$ and higher values of $\delta$ when $\Rv < 0$).

\fref{relspeed} also shows that when $\rho$ is fixed, relative speed depends on underlying epidemiological conditions---specifically, the absolute strength of the two strains. 
For example, even when the generation-interval distributions are identical ($\Gv=\Gw$, in this case), changing $\Rw$ from 0.49 (\fref{relspeed}, top-left panel) to 1.27 (\fref{relspeed}, bottom-right panel)---and thus $\Rv$ from 0.79 to 2.04---changes the relative speed $\delta$ from 0.08 to 0.11 when $\kappa=0.2$ (and from 0.06 to 0.14 when $\kappa=0.5$).
Differences in the generation-interval distributions exaggerate these changes.
Therefore, characterizing changes in the proportion of variants by assuming a fixed relative speed (e.g., by fitting a standard logistic growth curve \citep{switzerland2021variant} and using the resulting value of $\delta$) should be done with care.

@@

Instead of inferring relative strength from relative speed, one can separately estimate time-varying reproduction numbers $\RR(t)$ of the variant and the wild type from observed incidence and assumptions about the generation intervals, and then calculate the ratio---such methods have been used in previous analyses of the Alpha \citep{volz2021transmission}, Delta \citep{abbott2021estimating}, and Omicron variants \citep{pearson2021bounding}.
Broadly, there are two types of time-varying reproduction numbers: case reproduction number and instantaneous reproduction number.
The case reproduction number is defined as the average number of secondary infections caused by an individual infected at time $t$ and therefore depends on transmission after time $t$ \citep{wallinga2004different}.
The instantaneous reproduction number is defined as the average number of secondary infections that would be caused by an individual infected at time $t$ if conditions were to remain the same \citep{fraser2007estimating}; 
therefore, the instantaneous reproduction number only depends on transmission at time $t$ and is most appropriate for real-time evaluation of changes in transmission \citep{gostic2020practical}.
The estimation of the instantaneous reproduction number was popularized by \cite{cori2013new} and has been widely adopted in epidemiological analyses of SARS-CoV-2 \citep{abbott2020estimating,knight2020estimating,flaxman2020Rt,brauner2021inferring,li2021temporal}.

@@

When the wild type and the variant have the same (known) generation-interval distributions, the assumed generation-interval distribution matches the true distribution (\fref{Rtbias}D--F).
In this case, the estimated reproduction numbers match the true values (\fref{Rtbias}D); thus, their estimated ratio $\hat{\rho}(t)$ remains constant and $\hat{\RR}_{\textrm{var}}(t)=\Rv(t)$ (\fref{Rtbias}E).
However, when the true generation intervals of the variant have a shorter mean than the assumed distribution, we over-estimate $\Rv(t)$ during the growth phase and under-estimate $\Rv(t)$ during the decay phase (\fref{Rtbias}A);
therefore, the estimated ratio $\hat{\rho}(t)$ changes over time (\fref{Rtbias}B).
Conversely, when the true generation intervals of the variant have a longer mean than the assumed distribution, we find similar biases in opposite directions (\fref{Rtbias}G--H).

In practice, estimates of instantaneous reproduction numbers $\RR(t)$ (and therefore, their ratios) can be noisy due to limited data availability or model assumptions;
instead, we might want to estimate a single value of relative strength $\rho$.
For example, we can estimate $\rho$ by plotting the estimated strength of the variant $\hat{\RR}_{\textrm{var}}(t)$ against the estimated strength of the wild type $\hat{\RR}_{\textrm{wt}}(t)$---as presented in Figure 2 of \cite{volz2021transmission}---and performing a linear regression (\fref{Rtbias}C,F,I).
Here, we estimate the slope while fixing the intercept to zero: the regression line should go through the origin in theory because $\Rv = 0$ when $\Rw = 0$.
If $\rho$ is constant, and generation-interval distributions are correctly specified, we obtain a straight line with a slope of $\rho$ and intercept at zero (\fref{Rtbias}F).
However, when the assumed mean generation interval is longer than the that of the variant (\fref{Rtbias}C), we over-estimate the slope (and conversely, \fref{Rtbias}I).

@@

While detailed contact tracing data can help narrow down uncertainties in the generation-interval distributions, there are additional complexities to comparing generation-interval estimates of different variants.
The generation intervals can be measured in two different ways: backward and forward \citep{champredon2015intrinsic}.
The backward distribution starts from the cohort of infectees who were infected at the same time and looks at when their infectors were infected. 
The backward distribution is subject to dynamical biases because we are more likely to observe recent infections (and therefore shorter generation intervals) when the epidemic is growing.
On the other hand, the forward distribution starts from the cohort of infectors who were infected at the same time and looks at when they transmitted infections.
While the forward distribution is more reliable, changes in epidemic dynamics due to intervention measures and susceptible depletions can affect the shape of the distribution \citep{ali2020serial, nishiura2010time}.
For example, comparing generation-interval estimates from different intervention periods will necessarily bias the true differences in the generation-interval distributions of different variants.
